Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-16 paper.qmd
Added video links to the figure titles

```{r}
#| label: best-fit-tsne
#| fig-cap: "Lifting the fitted model into \\pD{}. Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to \\fD{} (b, c). The fit is reasonbly tight with the data, although it does not fully spread the full width. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice. Video of the langevitour animation is available at <>."
#| fig-width: 8
#| fig-height: 2

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 10
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP. The data has five Gaussian cluster in \\fD{}. The bottom row of plots shows a \\gD{} projection from a tour on \\fD{} revealing the differences generated by the layouts on the model fits.  We would expect the model fit to be like that in (a2) where it is distinctly is separate for each cluster but like a hairball in each. This would indicate the distinct clusters, each being fully \\fD{}. With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes.  Videos of the langevitour animations are available at <https://youtu.be/I-kxCwVfqiQ>, <https://youtu.be/gD1P01FUPyU>, and <https://youtu.be/MxJ_srOFQNk> respectively."

>>>>

```{r}
#| label: best-fit-tsne
#| fig-cap: "Lifting the fitted model into \\pD{}. Starting from a default tSNE layout (a) of nonlinear clusters data ($n =  2000$ and $p = 4$), from which the \\gD{} wireframe is constructed (a) and then lifted to \\fD{} (b, c). The fit is reasonbly tight with the data, although it does not fully spread the full width. This highlights a key characteristic of tSNE, which compresses the data when applying tSNE. Furthermore, the sparse space observed in the middle of the tSNE layout is a result of (hyper-)~parameter choice. The link to the video of the langevitour animation is available in @tbl-links."
#| fig-width: 8
#| fig-height: 2

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 10
#| fig-width: 15
#| fig-pos: H
#| fig-cap: "NLDR's organise points in the \\gD{} layout in different ways, possibly misleadingly, illustrated using three layouts: (a) tSNE, (b) UMAP, (c) PaCMAP. The data has five Gaussian cluster in \\fD{}. The bottom row of plots shows a \\gD{} projection from a tour on \\fD{} revealing the differences generated by the layouts on the model fits.  We would expect the model fit to be like that in (a2) where it is distinctly is separate for each cluster but like a hairball in each. This would indicate the distinct clusters, each being fully \\fD{}. With (c2), the curiousity is that the model is a \\gD{} pancake shape in \\fD{}, indicating that there is some ordering of points done by PaCMAP, posisbly along some principal component axes. The links to the videos of the langevitour animations are available in @tbl-links."
